Collective behavior of Janus Particles Suspended in a Viscous Fluid

### 2022-08-04 doc/PRF2022/PRF_main.tex
new Fig 0, updated text, updated Supplementary.tex

\section{Model parameters}
The interactions generated by the HAP-mobility problem formulation
\eqref{eq:stokes}--\eqref{eq:force}
lead to particle self-assembly for a broad range of parameters and
boundary conditions.  Generally speaking, the effective distance of the
hydrophobic interaction is set by the screening length $\rho$.
Smaller values of the repulsion distance $\rho_0$ decrease the distance
where attraction and repulsion are in balance and lead to more compact
particle assemblies. 
The rate of self assembly is proportional to the interfacial tension $\gamma$,
and inversely proportional to solvent viscosity $\mu$; it is
roughly inversely proportional to screening length $\rho.$

@@

@@

We modify $S_2$ to account for the bilayer and multilamellar structures.
In these cases, the directors are more or less uniformly distributed
along a circle, even though there is orientational order between
neighboring particles. To account for local order, we instead use
\begin{equation}
  \label{eq:localS2}
\tilde{S}_2(t) = \frac{1}{N_b} \sum_{i=1}^{N_b}
\frac{1}{2}(3\cos^2(\theta_i - \bar \theta_i) - 1),
\end{equation}
where $\bar \theta_i$ is the circular mean restricted to particles
indexed by $j$ with $\|\aa_i - \aa_j\| < 4c$. In practice, if a particle
is isolated and has no neighbors within a distance $4c$, then we exclude
it from the sum~\eqref{eq:localS2}.

@@

\subsection{Bilayer configurations (BC (i)) in a linear shear flow}
To begin, we place the bilayer structure (with BC (i)) from the inset of
Figure~\ref{fig:relax}(b) in a linear shear flow
\begin{equation}
\uu_\infty(\xx) = \dot\gamma \text{ ns }^{-1} ({\bf e}_y \cdot \mathbf{x}) {\bf e}_x,
\end{equation}
%
where $\dot\gamma$ is the dimensionless shear rate, and ${\ee}_x$ and
${\ee}_y$ are horizontal and vertical unit vectors, respectively. The
initial bilayer structure consists of a small self-enclosing bilayer
(vesicle) and several pieces of bilayers.
Figure~\ref{fig:BC1_shear}(a)--(c) shows a snapshot at $t = 0.5$
$\upmu$s of the bilayer for different shear rates $\dot\gamma=
0.05,0.075,0.01$. The dynamics of the bilayer can be found in the
supplementary movies, which show that the initial pieces of bilayers
merge and form a worm-like shape under a linear shear flow. At a low
shear rate, the bilayer constantly goes through rotation and extensional
deformation with the whole bilayer remaining intact
(Figure~\ref{fig:BC1_shear}(a)). As the shear rate increases to a
moderate value (Figure~\ref{fig:BC1_shear}(b)), the bilayer is observed
to shed a small piece of bilayer, similar to the asymmetric breakup of a
viscous drop in confinement \cite{DuFuZhuMaLi2016_AICHEJ}.
%\todo[inline]{missing citation}. 
In Figure~\ref{fig:BC1_shear}(c) the high shear rate gives rise to two
separate, nearly equal bilayers, similar to the symmetric drop breakup
under a linear shear flow \cite{Stone1994_ARFM}. 
%\todo[inline]{missing citation}.
%%
%\begin{figure}
%  \begin{center}
%    \includegraphics[width=1.0\textwidth]{Figures/Figure5.pdf}    
%  \end{center}
%  \vspace{-20pt}  
%  \caption{\label{fig:BC1_TG} Multiple component bilayers in a Taylor-Green when $V_0 = \{0.1, 0.15, 0.2\}$ at $t=0.2\ \mu$s. The pre-relaxed initial configuration is shown in inset of panel (a). 
% The streamlines are plotted in the background. 
%Panel (d) plots the free energies; panel (e) shows orientational parameter $\tilde{S}_2$; panel (f) shows positional parameter E.
%  }
%\end{figure}

We observe that the order parameter $\tilde S_2$ in
Figure~\ref{fig:BC1_shear}(e) is on average increasing. In the initial
configuration, the structure has three or four components.
%
Under the shear flow, the shedding and merging of bilayers results in an
overall decrease in distinct bilayer structures. Similarly, the free
energies are slightly decreasing (Figure~\ref{fig:BC1_shear}(d)), with
the overall changes occurring in the range of a few $\mathrm{k_BT}$/nm.
This suggests that the shear flow changes the energy landscape to move
the local equilibrium obtained from a random initial condition to an
alternate local equilibrium with less free energy and greater
orientational order.  Note that the free energy $F$ in
Figure~\ref{fig:BC1_shear}(d) has the unit of force ($\mathrm{k_BT}$/nm $\approx$ 4.114 pN)
because our system
is two dimensional.
%
%\begin{figure}
%  \begin{center}
%    \includegraphics[width=1.0\textwidth]{Figures/Figure6.pdf}        
%  \end{center}
%\caption{\label{fig:ves_TG} Single vesicle in a TG flow. Panels (a)-(c) are snapshots for $V_0=\{0.1, 0.15, 0.2\}$ at $t=0.6\ \mu$s where the pre-relaxed initial configuration is shown in inset of panel (a). The streamlines are plotted in the background.
%Panel (d) plots the free energies; panel (e) shows orientational parameter $\tilde{S}_2$; panel (f) shows positional parameter E.
%}
%\end{figure}
%
Finally, the strain parameter diverges in the large shear rate cases
(Figure~\ref{fig:BC1_shear}(f))
because the components eventually completely separate and move away from one another
under the shear flow.

@@

At shear rates $\dot\gamma=0.05, 0.066, 0.075$, the vesicle undergoes
familiar hydrodynamics, such as elongation along the extensional axis
and tank-treading motion, as shown in
Figure~\ref{fig:Ves_shear}(a)--(c).
%
The circular shape is also a local equilibrium, with somewhat less
energy than the disordered state c.f., Figure~\ref{fig:BC1_shear}(d) and
Figure~\ref{fig:Ves_shear}(d). But unlike the dynamics of bilayers in
Figure~\ref{fig:BC1_shear}, the energies of the JP vesicle in
Figure~\ref{fig:Ves_shear}(d) jump from the baseline value 30
$\mathrm{k_BT}$/nm to about 32 $\mathrm{k_BT}$/nm, signaling the rupture
of the vesicle into disconnected bilayers (see
Figure~\ref{fig:Ves_shear}(d), long and short dashed curves).  After the
rupture, each piece circles around the vorticity at the center without
reconnection while the energies fluctuate slightly around a constant.

@@

There is no appreciable pattern in the deformation for the
multicomponent bilayer case in TG flow. Figure~\ref{fig:BC1_TG}(a)--(c)
are the snapshots when $V_0=0.1,0.15,0.2$ at $t = 0.2$ $\upmu$s. At all
flow strengths, the bilayer components are separated and mixed by the
flow in neighboring cells. In the $V_0 = 0.2$ case, the flow breaks off
components consisting of only a few particles, leading an increase in
energy (Figure~\ref{fig:BC1_TG}(d), long dashed curve). There is no
apparente structure to the change in the orientational order. There is an initial
increase in the strain parameter due to component mixing, but the
strains remains bounded since the particles stay confined to the
neighbors of the central cell.

Patterns in the deformation, however, do emerge in the case of a vesicle
in TG flow. For a vesicle in a Taylor-Green flow,
Figure~\ref{fig:ves_TG}(a)--(c) are configurations at $t=0.2\ \upmu$s
where $V_0=0.1,0.15,0.2$. Here, the circular vesicle takes on a steady
square shape for moderate flow rates $V_0 = 0.1$ and $V_0 = 0.15$. In
these cases, there is little change in the vesicle free energy and order
(Figure~\ref{fig:ves_TG}(d)--(f)). The vesicle disintegrates for $V_0 =
0.2$, leading to a jump in energy, drop in order, and increase in
strain, suggesting that the vesicle has a critical TG flow strength in
the interval $0.15 < V_0 <  0.2$ separating the intact and ruptured
end-states.

@@

\subsection{Multilamellar and striated configurations in shear and TG flow}
We have simulated particles for the multilamellar (BC (ii)) and striated
(BC (iii)) boundary conditions, in both shear and TG background flow.
The results are qualitatively similar and will be presented in tandem.
For the multilamellar case, we place a 60-particle configuration (inset
of Figure~\ref{fig:BC2_shear}(a)) in shear flows with $\dot\gamma=0.05,
0.1, 0.15$ and in TG flow with flow rates $V_0=0.1, 0.15, 0.2$. For the
striated case, we also place a 60-particle configuration (inset of
Figure~\ref{fig:BC3_shear}(a)) in shear and TG background flows. Here,
somewhat larger rates are required to see appreciable deformations for
this boundary condition: $\dot\gamma=0.1, 0.125, 0.15$ and $V_0=0.2,
0.25, 0.3$, respectively.

@@

While increased flow strengths generally injected energy into the
suspension, the simulation results show that the unstructured bilayer
actually increases orientational order and has somewhat lower free
energy under low shear rates (Figure~\ref{fig:BC1_shear}(d,f)). This
effect suggests that in amphiphilic suspensions, it might be possible to
decrease the excess area of the system by subjecting the suspension to a
moderate shear flow. Not all background flows produce this
``organizing'' effect, as it is not observed in the TG background flow
case (Figure~\ref{fig:BC1_TG}). \citet{PhysRevLett.128.256102} have
shown how to drive particle clusters into arbitrary target shapes by
solving a first passage problem.

>>>>

\section{Model parameters and boundary conditions}
The interactions generated by the HAP-mobility problem formulation
\eqref{eq:stokes}--\eqref{eq:force}
lead to particle self-assembly for a broad range of parameters and
boundary conditions.  Generally speaking, the effective distance of the
hydrophobic interaction is set by the screening length $\rho$.
Smaller values of the repulsion distance $\rho_0$ decrease the distance
where attraction and repulsion are in balance and lead to more compact
particle assemblies. 
The rate of self assembly is proportional to the interfacial tension $\gamma$,
and inversely proportional to solvent viscosity $\mu$; it is
roughly inversely proportional to screening length $\rho.$

@@

\begin{figure}[h!]
\begin{center}
  \includegraphics[width=\textwidth]{Figures/Figure0.pdf}
\end{center}
\begin{caption}{\label{fig:boundary_conditions}
    The leftmost diagram illustrates the particle
    $\Gamma_i$ with center $\aa_i$, radius $c$, and director $\dd$,
    along with the angle $\alpha$ used to define $g(\xx)$ in \eqref{eq:bc-type}.
    The three righ plots show $g(\xx)$ (black curve) for BC (i), (ii), and (iii)
    when $\dd = (1,0)$. The surfaces are the respective solutions
    $u(\xx)$ of \eqref{eq:SL} for a single, isolated particle.  
  }
\end{caption}
\end{figure}

@@

We modify $S_2$ to account for the bilayer and multilamellar structures.
In these cases, the directors are more or less uniformly distributed
along a circle, even though there is orientational order between
neighboring particles. To account for local order, we instead use
\begin{equation}
  \label{eq:localS2}
\tilde{S}_2(t) = \frac{1}{N_b} \sum_{i=1}^{N_b}
\frac{1}{2}(3\cos^2(\theta_i - \bar \theta_i) - 1),
\end{equation}
where $\bar \theta_i$ is the circular mean restricted to particles
indexed by $j$ with $\|\aa_i - \aa_j\| < 4c$.
This cutoff distance was chosen empirically
so that the average includes nearest neighbors
but excludes next nearest neighbors.
In practice, if a particle
is isolated and has no neighbors within a distance $4c$, then we exclude
it from the sum~\eqref{eq:localS2}.

@@

\subsection{Bilayer configurations (BC (i)) in a linear shear flow}
To begin, we place the bilayer structure (with BC (i)) from the inset of
Figure~\ref{fig:relax}(b) in a linear shear flow
\begin{equation}
\uu_\infty(\xx) = \dot\gamma \text{ ns }^{-1} ({\bf e}_y \cdot \mathbf{x}) {\bf e}_x,
\end{equation}
%
where $\dot\gamma$ is the dimensionless shear rate, and ${\ee}_x$ and
${\ee}_y$ are horizontal and vertical unit vectors, respectively. The
initial bilayer structure consists of a small self-enclosing bilayer
(vesicle) and several pieces of bilayers.
Figure~\ref{fig:BC1_shear}(a)--(c) shows a snapshot at $t = 0.5$
$\upmu$s of the bilayer for different shear rates $\dot\gamma=
0.05,0.075,0.01$. The dynamics of the bilayer can be found in
Supplementary Movie S2, which shows
that the initial pieces of bilayers
merge and form a worm-like shape under a linear shear flow. At a low
shear rate, the bilayer constantly goes through rotation and extensional
deformation with the whole bilayer remaining intact
(Figure~\ref{fig:BC1_shear}(a)). As the shear rate increases to a
moderate value (Figure~\ref{fig:BC1_shear}(b)), the bilayer is observed
to shed a small piece of bilayer, similar to the asymmetric breakup of a
viscous drop in confinement \cite{DuFuZhuMaLi2016_AICHEJ}.
%\todo[inline]{missing citation}. 
In Figure~\ref{fig:BC1_shear}(c) the high shear rate gives rise to two
separate, nearly equal bilayers, similar to the symmetric drop breakup
under a linear shear flow \cite{Stone1994_ARFM}. 
Refer to Supplementary Movie S4.

%\todo[inline]{missing citation}.
%%
%\begin{figure}
%  \begin{center}
%    \includegraphics[width=1.0\textwidth]{Figures/Figure5.pdf}    
%  \end{center}
%  \vspace{-20pt}  
%  \caption{\label{fig:BC1_TG} Multiple component bilayers in a Taylor-Green when $V_0 = \{0.1, 0.15, 0.2\}$ at $t=0.2\ \mu$s. The pre-relaxed initial configuration is shown in inset of panel (a). 
% The streamlines are plotted in the background. 
%Panel (d) plots the free energies; panel (e) shows orientational parameter $\tilde{S}_2$; panel (f) shows positional parameter E.
%  }
%\end{figure}

We observe that the order parameter $\tilde S_2$ in
Figure~\ref{fig:BC1_shear}(e) is on average increasing. In the initial
configuration, the structure has three or four components.
%
Under the shear flow, the shedding and merging of bilayers results in an
overall decrease in distinct bilayer structures. Similarly, the free
energies are slightly decreasing (Figure~\ref{fig:BC1_shear}(d)), with
the overall changes occurring in the range of a few $\mathrm{k_BT}$/nm.
This suggests that the shear flow changes the energy landscape to move
the local equilibrium obtained from a random initial condition to an
alternate local equilibrium with less free energy and greater
orientational order  (Supplementary Movie S2).  Note that the free energy $F$ in
Figure~\ref{fig:BC1_shear}(d) has the unit of force ($\mathrm{k_BT}$/nm $\approx$ 4.114 pN)
because our system
is two dimensional.
%
%\begin{figure}
%  \begin{center}
%    \includegraphics[width=1.0\textwidth]{Figures/Figure6.pdf}        
%  \end{center}
%\caption{\label{fig:ves_TG} Single vesicle in a TG flow. Panels (a)-(c) are snapshots for $V_0=\{0.1, 0.15, 0.2\}$ at $t=0.6\ \mu$s where the pre-relaxed initial configuration is shown in inset of panel (a). The streamlines are plotted in the background.
%Panel (d) plots the free energies; panel (e) shows orientational parameter $\tilde{S}_2$; panel (f) shows positional parameter E.
%}
%\end{figure}
%
Finally, the strain parameter diverges in the large shear rate cases
(Figure~\ref{fig:BC1_shear}(f))
because the components eventually completely separate and move away from one another
under the shear flow.

@@

At shear rates $\dot\gamma=0.05, 0.066, 0.075$, the vesicle undergoes
familiar hydrodynamics, such as elongation along the extensional axis
and tank-treading motion, as shown in
Figure~\ref{fig:Ves_shear}(a)--(c)
and Supplementary Movie S2.
%
The circular shape is also a local equilibrium, with somewhat less
energy than the disordered state c.f., Figure~\ref{fig:BC1_shear}(d) and
Figure~\ref{fig:Ves_shear}(d). But unlike the dynamics of bilayers in
Figure~\ref{fig:BC1_shear}, the energies of the JP vesicle in
Figure~\ref{fig:Ves_shear}(d) jump from the baseline value 30
$\mathrm{k_BT}$/nm to about 32 $\mathrm{k_BT}$/nm, signaling the rupture
of the vesicle into disconnected bilayers (see
Figure~\ref{fig:Ves_shear}(d), long and short dashed curves).  After the
rupture, each piece circles around the vorticity at the center without
reconnection while the energies fluctuate slightly around a constant.

@@

There is no appreciable pattern in the deformation for the
multicomponent bilayer case in TG flow
(Supplementary Movie S3, right panel).
Figure~\ref{fig:BC1_TG}(a)--(c)
are the snapshots when $V_0=0.1,0.15,0.2$ at $t = 0.2$ $\upmu$s. At all
flow strengths, the bilayer components are separated and mixed by the
flow in neighboring cells. In the $V_0 = 0.2$ case, the flow breaks off
components consisting of only a few particles, leading an increase in
energy (Figure~\ref{fig:BC1_TG}(d), long dashed curve). There is no
apparente structure to the change in the orientational order. There is an initial
increase in the strain parameter due to component mixing, but the
strains remains bounded since the particles stay confined to the
neighbors of the central cell.

Patterns in the deformation, however, do emerge in the case of a vesicle
in TG flow (Supplementary Movie S3, left panel). For a vesicle in a Taylor-Green flow,
Figure~\ref{fig:ves_TG}(a)--(c) are configurations at $t=0.2\ \upmu$s
where $V_0=0.1,0.15,0.2$. Here, the circular vesicle takes on a steady
square shape for moderate flow rates $V_0 = 0.1$ and $V_0 = 0.15$. In
these cases, there is little change in the vesicle free energy and order
(Figure~\ref{fig:ves_TG}(d)--(f)). The vesicle disintegrates for $V_0 =
0.2$ (Supplementary Movie S5, left panel), leading to a jump in energy, drop in order, and increase in
strain, suggesting that the vesicle has a critical TG flow strength in
the interval $0.15 < V_0 <  0.2$ separating the intact and ruptured
end-states.

@@

\subsection{Multilamellar and striated configurations in shear and TG flow}
We have simulated particles for the multilamellar (BC (ii)) and striated
(BC (iii)) boundary conditions, in both shear and TG background flow.
The results are qualitatively similar and will be presented in tandem.
For the multilamellar case, we place a 60-particle configuration (inset
of Figure~\ref{fig:BC2_shear}(a)) in shear flows with $\dot\gamma=0.05,
0.1, 0.15$ and in TG flow with flow rates $V_0=0.1, 0.15, 0.2$. For the
striated case, we also place a 60-particle configuration (inset of
Figure~\ref{fig:BC3_shear}(a)) in shear and TG background flows. Here,
somewhat larger rates are required to see appreciable deformations for
this boundary condition: $\dot\gamma=0.1, 0.125, 0.15$ and $V_0=0.2,
0.25, 0.3$, respectively.
The BC (ii) and BC (iii) in shear flow cases are visualized in
the latter half of 
Supplementary Movie S2 (low shear rate) and 
Supplementary Movie S4 (high shear rate), respectively.
The low and high flow rates for TG flow are shown in
the latter half of
Supplementary Movie S3 and 
Supplementary Movie S5, respectively.

@@

While increased flow strengths generally injected energy into the
suspension, the simulation results show that the unstructured bilayer
actually increases orientational order and has somewhat lower free
energy under low shear rates (Figure~\ref{fig:BC1_shear}(d,f)). This
effect suggests that in amphiphilic suspensions, it might be possible to
decrease the excess area of the system by subjecting the suspension to a
moderate shear flow (Supplementary Movie S2, right panel).
Not all background flows produce this
``organizing'' effect, as it is not observed in the TG background flow
case (Figure~\ref{fig:BC1_TG}, Supplementary Movie S3, right panel).
\citet{PhysRevLett.128.256102} have
shown how to drive particle clusters into arbitrary target shapes by
solving a first passage problem.
